An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-23 VLDB2026/main.tex
1

@@

	 %\underline{Faiss generates slightly smaller indexes due to its efficient index design}.

	\textit{\textbf{Build peak memory.}}
	As shown in Figure~\ref{fig:build_memory_mb_comparison}, 
	%on small-scale datasets, all methods except NHQ exhibit low memory usage. 
	\textcolor{orange}{NHQ has the highest peak memory usage on almost all datasets.}
	This is because NHQ adopts the kgraph approach for index construction, which requires loading the entire graph data into memory and maintaining a complete graph structure, leading to relatively high memory overhead.

	On large-scale datasets, CAPS exhibits slightly lower memory usage compared to other algorithms. \textcolor{orange}{The reason is that CAPS avoids redundant storage and graph structure overhead through hierarchical partitioning and AFT, and achieves a compact, sparse and controllable index. }.

	\textit{\textbf{ \textcolor{orange}{Search peak memory.}}}
	 \textcolor{orange}{As shown in Figure~\ref{fig:search_memory_mb_comparison}, we can find that the peak memory usage of each algorithm during search does not differ significantly. Among them, CAPS and NHQ have relatively higher peak memory usage, while the peak memory usage of database is comparatively lower.}

	\subsubsection{Performance Evaluation}

@@

	\textit{Effect of Multi-Attribute Building on Algorithm Performance (M vs. S).}
	As shown in Figure \ref{fig:exp_2_1}, \textcolor{red}{Puck demonstrates nearly no performance degradation, because it accurately identifies the nearest clusters during a search, keeping the number of candidates constant, no matter how many attributes were used to build the index}. CAPS shows a slight performance decline due to the increased query overhead caused by excessive attribute-based grouping. In contrast, the performance of StitchedVamana, FilteredVamana, NHQ, and UNG drops significantly, with the following reasons: 1) FilteredVamana increases the complexity introduced by retaining attribute during index construction. 2) The complexity of StitchedVamana increases because data points may duplicate across multiple subgraphs. 3) The fusion distance of NHQ exhibits sensitivity to the number of attributes, which degrades its effectiveness. 4) UNG has poor performance because large number of entry points that need to be traversed. It is difficult for a graph with multiple entries to have good connectivity, degenerating into a situation closer to inverted retrieval, and the effect may be greatly impaired.

@@

	(1) \textit{Algorithms optimized for low AS.}  
	This group includes ACORN-1, Faiss+HQI\_Batch, StitchedVamana, UNG, Faiss, Puck, and CAPS, as illustrated in Figure~\ref{fig:exp_5_2_1}.

>>>>

	for big dataset:

@@

	 for big dataset:
	 %\underline{Faiss generates slightly smaller indexes due to its efficient index design}.

	\textit{\textbf{Build peak memory.}}
%	As shown in Figure~\ref{fig:build_memory_mb_comparison}, 
%	%on small-scale datasets, all methods except NHQ exhibit low memory usage. 
%	\textcolor{orange}{NHQ has the highest peak memory usage on almost all datasets.}
%	This is because NHQ adopts the kgraph approach for index construction, which requires loading the entire graph data into memory and maintaining a complete graph structure, leading to relatively high memory overhead. 
%	
%	On large-scale datasets, CAPS exhibits slightly lower memory usage compared to other algorithms. \textcolor{orange}{The reason is that CAPS avoids redundant storage and graph structure overhead through hierarchical partitioning and AFT, and achieves a compact, sparse and controllable index. }.
	\textcolor{orange}{As shown in the Figure~\ref{fig:build_memory_mb_comparison}, the build peak memory usage is generally correlated with the sum of the index size generated by the algorithm and the dataset size. This trend holds consistently across different datasets. However, PASE breaks this pattern due to its on-demand loading strategy, resulting in peak memory usage even lower than the index size itself. For large-scale datasets, algorithms that produce larger indexes tend to cause significantly higher peak memory consumption during construction.}

	\textit{\textbf{ \textcolor{orange}{Search peak memory.}}}
%	 \textcolor{orange}{As shown in Figure~\ref{fig:search_memory_mb_comparison}, we can find that the peak memory usage of each algorithm during search does not differ significantly. Among them, CAPS and NHQ have relatively higher peak memory usage, while the peak memory usage of database is comparatively lower.}
	\textcolor{orange}{For most algorithms, the query phase requires loading both the index and the dataset into memory, so its peak memory usage is usually comparable to that of the construction phase. However, as shown in the Figure~\ref{fig:search_memory_mb_comparison}, CAPS exhibits significantly higher peak memory during querying due to the need to dynamically maintain large candidate sets. In contrast, Milvus has lower query-time memory usage because it does not fully load the index and incurs additional metadata overhead during construction.
	}
	\subsubsection{Performance Evaluation}

@@

	\textit{Effect of Multi-Attribute Building on Algorithm Performance (M vs. S).}
	As shown in Figure \ref{fig:exp_2_1}, Puck demonstrates nearly no performance degradation.  CAPS shows a slight performance decline due to the increased query overhead caused by excessive attribute-based grouping. In contrast, the performance of StitchedVamana, FilteredVamana, NHQ, and UNG drops significantly, with the following reasons: 1) FilteredVamana increases the complexity introduced by retaining attribute during index construction. 2) The complexity of StitchedVamana increases because data points may duplicate across multiple subgraphs. 3) The fusion distance of NHQ exhibits sensitivity to the number of attributes, which degrades its effectiveness. 4) UNG has poor performance because large number of entry points that need to be traversed. It is difficult for a graph with multiple entries to have good connectivity, degenerating into a situation closer to inverted retrieval, and the effect may be greatly impaired.

@@

	(1) \textit{Algorithms optimized for low AS.}  
	This group includes ACORN-1, CAPS, StitchedVamana, UNG, Puck, Faiss, and Faiss+HQI\_Batch as illustrated in Figure~\ref{fig:exp_5_2_1}.
